---
title: "415 Project"
output:
  pdf_document: default
---

#4 Data Import & Cleaning
###4.1 Data Import & Cleaning:
```{r}
library(tidyverse)
library(readr)
library(ggpubr)
gym_data <- read.csv("gym_data.csv", head = TRUE, sep = ",")
problems(gym_data)
```

###4.2 Data Cleaning 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Creating columns for year and day from the date column
gym_data <- gym_data %>% mutate(day = substr(date, 9, 10)) %>% mutate(year = substr(date, 1, 4))
# Assert type of year and day to be numeric
typeof(gym_data$year) 
typeof(gym_data$day)
gym_data$year<-as.numeric(gym_data$year)
gym_data$day <- as.numeric(gym_data$day)
#Getting the observations between Nov. 2016 and Feb. 2017
data_set <- gym_data[c(50466:60861),c(1:13)]
data_set<-data_set%>%select(timestamp,number_people,day_of_week,is_weekend,is_holiday,temperature,is_start_of_semester,is_during_semester,month,hour,day,year)
```

#5 Variation of Single Variables
###5.1 number_people and temperature
```{r, echo=FALSE, message=FALSE, warning=FALSE}
np<-ggplot(data_set)+ geom_histogram(aes(x = number_people))+theme_bw()+
  ggtitle("Variable: number_people")+
  xlab("number of people in the gym") 
 # xlim(2,max(data_set$number_people))
#summary(data_set$number_people)

t<-ggplot(data_set)+ geom_histogram(aes(x = temperature))+theme_bw()+
  ggtitle("Variable: temperature")+
  xlab("temperature of the day") 

#hist(data_set$temperature)
#summary(data_set$temperature)

ggarrange(np, t, ncol = 2)

```

The variable of number_people roughly normally distributed except that there are some values that are close to zero. This actually make sense for our dataset. Because our dataset records number of people every 10 minutes for 24 hours, which means it would contain observations that is 3:10AM in the morning and this would not be a common time that people go to gym. It has a median of 16 and mean of 16.07. Maximun value is 88 and minimun is 0.

<br />The variable of temperature roughly normally distributed and skewed to the right. It has a median of 52.33 fahrenheit and mean of 53.34 fahrenheit. Maximun value is 73.39 fahrenheit and minimun is 38.92 fahrenheit.


###5.2 day_of_week, is_weekend, is_holiday
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#unique(data_set$day_of_week)
#unique(data_set$is_weekend)
#unique(data_set$is_holiday)

pp1<-ggplot(data_set,mapping = aes(x=as.character(day_of_week),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("What day it is duriing the week")+
  ylab("Number of people")+
  ggtitle("number_people vs. day_of_week")

pp2<-ggplot(data_set,mapping = aes(x=as.character(is_weekend),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("Whether or not is during weekend")+
  ylab("Number of people")+
  ggtitle("number_people vs. is_weekend")

pp3<-ggplot(data_set,mapping = aes(x=as.character(is_holiday),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("Whether or not is during holidays")+
  ylab("Number of people")+
  ggtitle("number_people vs. is_holiday")

ggarrange(pp1,pp2,pp3,ncol = 3, nrow = 1)
```

"day_of_week" is a categorical variables. "is_weekend" and "is_holiday" are binary variables.
The bar graphs above show the distribution of our "day_of_week", "is_weekend", and "is_holiday" variables. As we can imagine, the "day_of_week" variable has a roughly uniform distribution; "is_weekend" variable's distribution is roughly about 5:2.


###5.3 month,hour,day
```{r, echo=FALSE, message=FALSE, warning=FALSE}

p1 <- ggplot(data_set,mapping = aes(x=as.character(month),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("Month")+
  ylab("Number of people")+
  ggtitle("number_people vs. month")

p2 <- ggplot(data_set,mapping = aes(x=as.character(hour),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("Hour")+
  ylab("Number of people")+
  ggtitle("number_people vs. hour") + coord_flip()

p3 <- ggplot(data_set,mapping = aes(x=as.character(day),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("day")+
  ylab("Number of people")+
  ggtitle("number_people vs. day") + coord_flip()

ggarrange(p1,p2,p3,ncol = 3,nrow=1)
#hist(data_set$month)
#summary(data_set$month)
#hist(data_set$hour)
#summary(data_set$hour)
#hist(as.numeric(data_set$day))
#summary(as.numeric(data_set$day))
```

###5.5 is_start_of_semester, is_during_semester
```{r, echo=FALSE, message=FALSE, warning=FALSE}
s<-ggplot(gym_data, mapping = aes(x=as.character(is_start_of_semester),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("Whether in the start of semester or not")+
  ylab("Number of people")+
  ggtitle("number_people vs. is_start_of_semester")

d<-ggplot(gym_data,mapping = aes(x=as.character(is_during_semester),y=number_people))+
  geom_boxplot() +
  theme_bw() +
  xlab("Whether in semester or not")+
  ylab("Number of people")+
  ggtitle("number_people vs. is_during_semester")

ggarrange(s, d, ncol = 2, nrow = 1)

```
"is_start_of_semester"" and "is_during_semester" are binary variables. As we would think, the "is_during_semester" variable has a distribution roughly about 1:2, since usually during the time of 2016-11-01 to 2017-02-28 we have more than a month of winter break.

#6 Covariation between Multiple Variables
<br>
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(GGally)
library(ggpubr)
library(corrplot)
d <- data_set # Abbreviation
d$date<-NULL
corrplot(cor(d), tl.cex=1,tl.col = "black")
```
A correlation matrix in R computes correlation majorly between continuous variables, but it also computes a corrected pearson's correlation between a mix of categorical and continuous variables. According to our correlation matrix of the dataframe, Timestamp, temperature, is_during_semester, and hour are relatively more correlated to crowdedness of the gym than other variables. Also, independent variable hour and timestamp, year and month seems to be closely correlated. This might incure the risk of collinearity while doing statistical modelling. Therefore we would focus on one of the two variables in the pair hour and timestamp, and pair year and month. Chosing the variable would be implemented in later analysis. As we mentioned, correlation computed for categorical and continuous variables might not be as accurate as the ones computed for continuous variables. Therefore, we need more evidence for the correlation between month and year, hour and timestamp. In this case, looking at vif gives us more information:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(usdm)
tes <- d
vif(d)
tes$year<-NULL
tes$hour<-NULL
vif(tes)
```

Year, month, timestamp, and hour all have a vif above 100, which is a strong indicator of collinearity. Since timestamp and month captures time variations in greater detail, we choose to delete hour and year from our dataset. vifs of variables in the new dataset all goes down to below 4, which is a acceptable range by convention. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
d %>% dplyr::select(timestamp, temperature,number_people) %>%
   ggpairs(mapping=ggplot2::aes(colour = as.character(d$month)))
```

For covariation within continuous variables, we can see that dependent variables have moderate correlations of around 0.5 with our independent variable, and dependent variables have low correlation of 0.19. More importantly, it seems like as timestamp or tenmperature goes up, the crowdedness of the gym increases as well. Coloring the trend with month, we can see that the temperature is highest in Nov. Also, Nov has the highest crowdedness possible. This may suggest an effect of temperature on the crowdedness of the gym. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

sub <- function(a){
  if(a==0){
    res2="P-value < 2e-16"
    return(res2)
  }
  else{
    a=as.character(a)
    a1=substr(a,1,4)
    a2=substr(a, nchar(a)-4+1, nchar(a))
    res1 = paste(a1,a2,sep="")
    res2=paste("P-value =", res1, sep=" ")
    return(res2)
  }
  
}
cor1 <- sub(summary(aov(d$number_people ~d$day_of_week))[[1]][["Pr(>F)"]][1])
dow<- d %>% group_by(day_of_week) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(day_of_week), number_people, FUN = median), y=number_people)) +
                    labs(x = "day_of_week") + geom_boxplot(aes(fill=count))+ggtitle(cor1)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

cor2<- sub(summary(aov(d$number_people ~d$is_weekend))[[1]][["Pr(>F)"]][1])
isw<-d %>% group_by(is_weekend) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(is_weekend), number_people, FUN = median), y=number_people)) +
                    labs(x = "is_weekend") + geom_boxplot(aes(fill=count)) +ggtitle(cor2)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

cor3<- sub(summary(aov(d$number_people ~d$is_holiday))[[1]][["Pr(>F)"]][1])
ish<-d %>% group_by(is_holiday) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(is_holiday), number_people, FUN = median), y=number_people)) +
                    labs(x = "is_holiday") + geom_boxplot(aes(fill=count)) +ggtitle(cor3)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

cor4<- sub(summary(aov(d$number_people ~d$is_start_of_semester))[[1]][["Pr(>F)"]][1])
issfs<-d %>% group_by(is_start_of_semester) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(is_start_of_semester), number_people, FUN = median), y=number_people)) +
                    labs(x = "is_start_of_semester") + geom_boxplot(aes(fill=count)) +ggtitle(cor4)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

cor5<- sub(summary(aov(d$number_people ~d$is_during_semester))[[1]][["Pr(>F)"]][1])
isds<-d %>% group_by(is_during_semester) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(is_during_semester), number_people, FUN = median), y=number_people)) +
                    labs(x = "is_during_semester") + geom_boxplot(aes(fill=count)) +ggtitle(cor5)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

cor6<- sub(summary(aov(d$number_people ~d$month))[[1]][["Pr(>F)"]][1])
mon <-d %>% group_by(month) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(month), number_people, FUN = median), y=number_people)) +
                    labs(x = "month") + geom_boxplot(aes(fill=count)) +ggtitle(cor6)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

cor7<- sub(summary(aov(d$number_people ~d$hour))[[1]][["Pr(>F)"]][1])
hr <-d %>% group_by(hour) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(hour), number_people, FUN = median), y=number_people)) +
                    labs(x = "hour") + geom_boxplot(aes(fill=count)) +ggtitle(cor7)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

cor8<- sub(summary(aov(d$number_people ~d$day))[[1]][["Pr(>F)"]][1])
day <-d %>% group_by(day) %>% 
                    mutate(count=n()) %>% 
                    ggplot(aes(x=reorder(factor(day), number_people, FUN = median), y=number_people)) +
                    labs(x = "day") + geom_boxplot(aes(fill=count)) +ggtitle(cor8)+theme(plot.title=element_text(size=10.5, face="bold.italic"))

ggarrange(dow,isw, ish, issfs, ncol = 2, nrow = 2)

```

Above each boxplot are the p-value for the anova analysis of statistical significance between two variables. A p-value below 0.05 by convention suggests statistical significance, or a strong correlation between two variables. As we can see from the boxplots, there is significant difference between crowdness in weekend and weekdays. The count for weekdays are significantly higher. Suggesting that the weekly routine for most people are during weekdays. The same situation applies to is_holidays. However, for during semester and start of semester. Both counts seems to be really high. The two boxplots seems to be contradictory, and are something we should investigate further into the analysis. Also, by looking at month, there seems to be less crowdedness at the end and the beginning of the year. This is probably due to the Chirstmas and winter holidays. 

```{r, echo = FALSE, message=FALSE,warning=FALSE}
ggarrange(isds, mon, ncol = 2, nrow = 1)
```

```{r, echo = FALSE, message=FALSE,warning=FALSE}
ggarrange(hr, day, ncol = 1, nrow = 2)

```

From boxplot for hours and day, we can see that crowdedness in gym are typically between 9 and 22. And there's significantly low activities in 23th, 24th, 25th, and 26th of each month.

# Modeling and Analysis

### Fit model
Recall that we have excluded hour and year from the pool of independent variables, the summary of our multiple linear regression are:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
mls <- lm(d$number_people ~ d$timestamp+d$day_of_week+d$is_weekend+d$is_holiday+d$temperature+d$is_start_of_semester+d$is_during_semester+d$month+d$day)
summary(mls)

```

Since is_weekend and day are statistically insignificant based on a significance level of 0.05, we exclude these two variables and fit the model again: 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
mls2 <- lm(d$number_people~ d$timestamp+d$day_of_week+d$is_holiday+d$temperature+d$is_start_of_semester+d$is_during_semester+d$month)
summary(mls2)
```
In mls2, we build an regression of number_people on timestamp, day_of_week, is_holiday, temperature, is_start_of_semester, is_during_semester, month. All the varibales are significient (all below 0.001). In this model, the betas of variables day_of_week, is_holiday and month are negative, which makes sense. When it gets closer to the weekends and holidays, people tends to go to gyms less and spend their time in other things. The month variable has an negative effects here can be interpret as that since we are only taking Jan, Feb, Nov and Dec, Nov and Dec are typical final exam period for school and students spend less time in the gym. Therefore, it makes sense to see there is an negtive for month varibale. For variables timestamp, temperature, is_start_of_semester and is_during_semester, they all have an coefficient larger than zero. They also make sense. Because more people tend to go to gym during evening after school/work; people tends to go to gym when the weather is warm; students typically have less school work and more time for gym in the start of the semester; for a college gym, there will be more people during the semester than not during the semester.

Now that all variables are statistically significant, for next step we consider a interaction term that we believe are statistically:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
mls2_interactions <- lm(d$number_people~d$timestamp*d$day_of_week+d$is_holiday+d$temperature+d$is_start_of_semester+d$is_during_semester+d$month)
summary(mls2_interactions)
```
### Diagnosis
A crucial step after fitting the model is to check the residual for assumptions:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot(mls2_interactions)
```

As we can see, the qq plot conveys that our standardized residuals are normally distributed as they follows a straight line while comparing with the theoretical quantiles. However, when we are taking a closer look at the plot with Residuals versus fitted values, we observed an very unusual line going downwards. We suspect this might conveys the same problem as the shown in the distribution of our response variable:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
qqnorm(d$number_people)
```
As shown in the plot above, a very large proportion of our data contains 0, which makes our qq plot incredibly tail heavy, and violates the normality assumption. The way of transforming the dataset is somewhat complicated and we shall discuss possible transformations/improvements to the dataset in the discussion section.

### Prediction
```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(modelr)
require(ggplot2)
d <- filter(d, hour > 5 | hour == 0)
g <- d%>% add_predictions(mls)
ggplot(g, aes(x= factor(month))) + # basic graphical object
geom_point(aes(y=pred), colour="red",position = "jitter", size=0.5) + # first layer
geom_point(aes(y=number_people), colour="green", position = "jitter",size=0.5) + # second layer
xlab("Month")+
ylab("Number of people")+
ggtitle("Observation and prediction")+
labs(caption = "Figure 25: Predictions based on models")

```


```{r}
library(modelr)
g <- d%>% add_predictions(mls)
g%>%ggplot(aes(x= factor(month), y= pred))+geom_point(position = "jitter")

g%>%ggplot(aes(x= factor(month), y= number_people))+geom_point(position = "jitter")
```
